The impact of forest management on the temperature sensitivity of organic matter turnover

### 2024-08-29 Manuscript_draft/Draft.Rmd
Some ideas on introduction

# Introduction
The soil is one of the largest terrestrial reservoirs of carbon, containing approximately three times more carbon than the atmosphere or the vegetation on Earth (IPCC, 2021). Soil carbon is released to the atmosphere as carbon dioxide (CO₂) through decomposition of organic matter by microbes (heterotrophic respiration) and root respiration (autotrophic respiration). This process, i.e., soil respiration, constitutes the main carbon efflux into the atmosphere (ref). Under the new climate change scenarios increases of CO₂ efflux from soils are expected (Schlesinger and Andrews, 2000), in turn this can imply a positive feedback to future warming (ref). Because soil respiration is highly sensitive to soil temperature and soil moisture (ref), understanding the responses of soil respiration to these microclimatic conditions, which are affected by global change factors as climate change and forest management is critical for predicting future atmospheric CO2 concentrations.

## Soil C turnover is controlled by temperature and moisture

## Forest management changes the soil C turnover sensitivity to microclimate

## Climate change will influence soil C turnover globally

In this study, during three years we have monthly recorded data on soil respiration and microclimate in three replicated Quercus forests across Europe subjected to different forest management strategies. We have analysed the sensitivity of soil respiration to microclimate (i.e., soil temperature and soil moisture) by constructing a mechanistic model…

>>>>

# Introduction
The soil is one of the largest terrestrial reservoirs of carbon, containing approximately three times more carbon than the atmosphere or the vegetation on Earth (IPCC, 2021). Soil carbon is released to the atmosphere as carbon dioxide (CO2) through decomposition of organic matter by microbes (heterotrophic respiration) and root respiration (autotrophic respiration). This process (i.e., soil respiration) constitutes the main carbon efflux into the atmosphere (ref). Under the new climate change scenarios increases of CO2 efflux from soils are expected (Schlesinger and Andrews, 2000), this can imply a positive feedback to future warming if stored carbon is mobilized and carbon input to soil does not increase (Lamberty&Thomson, 2010). Because soil respiration is highly sensitive to soil temperature and soil moisture (Nissan et al. 2022), understanding the responses of soil respiration to these microclimatic conditions, which are affected by global change factors as climate change and forest management, is critical for predicting future atmospheric CO2 concentrations. Temperature affects respiration rates positively (Lloyd and Taylor, 1994), by increasing the metabolic rates of soil microorganisms and enzyme efficiency (ref). However, the relationship with soil moisture is non-linear. Low water content limit nutrient diffusion and enzymes and exert drought-stress on microbes while high water contents limits oxygen (O2) diffusion, creating anaerobic conditions that suppress aerobic microbial respiration (ref). Something here about modelling?

Forest management have the potential to affect respiration rates by modifying the microclimate. Removal of trees by forest logging increases the exposure of soil to temperature fluctuations and alter the soil water content either by decreasing water through soil evaporation or by increasing it by reducing the transpiration of water by trees (Zhang et al 2018). The degree of the response of soil will be dependent on the severity of the forest intervention. Thus, more severe forestry intervention such as clear-cuts may increase soil moisture, enhance heat and frost damage (Kovaks et al 2019; Zhang et al 2018) and it seems to reduce respiration rates (Yang et el 2022; Zhang et al 2018). Less severe interventions such as thinning, have shown to increase respiration in broadleaf forests in a meta-analysis (Zhang et al 2018). Negative effects of forest management on soil could be buffered by the addition of forest residues on soil (i.e., slash addition), which can improve the microclimatic conditions, nutrient fertility and biodiversity (ref). We do not know whether the response of soil CO2 effluxes to the microclimatic conditions (i.e. soil temperature and moisture) differs among different forest management strategies. Understanding the sensitivity of respiration to microclimate under different forest management strategies is crucial in a context of climate-smart forestry (CSF) development.

Here, we hypothesized that logging of trees will decrease the respiration rates because of the strong decrease in vegetation activity and microbiota associated. In turn, addition of slash will provide labile organic matter to decompose increasing the respiration rates. We also hypothesized that the variation in soil respiration would be mainly dependent on microclimatic variables (soil temperature and moisture), especially in the management treatments where vegetation was removed, since heterotrophic respiration will prevail over autotrophic respiration. Moreover, the sensitivity of soil respiration to microclimatic conditions will depend on the type of climate/forest, being the Mediterranean the most sensitive as the microbial communities will be better adapted to the strong seasonal changes in water availability.

To test these hypotheses, we have monthly recorded data on soil respiration and microclimate in three replicated Quercus forests across Europe subjected to different forest management strategies during three years. We have analysed the sensitivity of soil respiration to microclimate (i.e., soil temperature and soil moisture) by constructing a mechanistic model…
